---
title: "ballots-returned-tracker-philly"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(readxl)
library(openxlsx)
library(stringr)

```

```{css, echo=FALSE}

h1.title {
  font-family: roboto;
  color: transparent;
  font-weight: 700;
  text-align: left;
  font-size: 12px;
  padding: 0px;
  height: 0px;
}

.footer {
  #font-family: roboto;
  color: black;
  text-align: left;
  font-size: 12px;
    padding: 5px;
  font-style: italic;
}

h1 {
  #font-family: roboto;
  color: black;
  font-weight: bolder;
  text-align: center;
  font-size: 36px;
  margin-top: 0;
  margin-bottom: 30px;
  
}
h2 {
  #font-family: roboto;
  font-weight: 500;
  color: black;
  text-align: center;
  font-size: 18px;
  margin-top: 0;
  margin-bottom: 2;
}
.bignumber {
  #font-family: roboto;
  color: white;
  font-weight: 900;
  text-align: center;
  font-size: 40px;
  line-height: 65px;
  height: 65px;
  margin-top: 0;
  margin-bottom: 0;
}
body {
  color: black;
  #font-family: roboto;
  font-weight: 400;
  font-size: 18px;
}
h4 {
  font-size: 14px;
  font-style: italic;
  text-align: center;
}
  
   * {
      box-sizing: border-box;
   }
   .card {
      color: white;
      float: left;
      width: calc(25%);
      padding: 5px;
      border-radius: 10px;
      margin-left: 50px;
      margin-right: 50px;
      margin-top: 25px;
      margin-bottom: 25px;
      height: 100%;
   }
   .card p {
     #font-family: roboto;
     text-align: center;
     font-size: 18px;
     margin-bottom: 0;
   }
   .cardContainer:after {
      content: "";
      display: block;
      clear: both;
   }
   
   @media screen and (max-width: 760px) {
      .bignumber {
         font-size: 36px;
      }
     
     .card p {
         font-size: 13px;
      }


   }
   
   @media screen and (max-width: 650px) {
      .card {
         width: 100%;
      }
      h1.title {
        font-size: 22px;
      }
      
      .bignumber {
         font-size: 35px;
      }
     
     .card p {
         font-size: 14px;
      }
      
      
   }
```

```{r, echo=FALSE, warning=FALSE}


```

```{r, echo=FALSE, warning=FALSE}

#counties <- c()

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

today_UTC <- as.POSIXct(Sys.time()) #get todays date as UTC
today_posix <- format(today_UTC, tz="America/New_York",usetz=TRUE)
today <- as.Date(substr(as.character(today_posix), 1,10))
yesterday <- as.Date(today-1) #get yesterday's date
filename <- paste(sep="", "vbm-statistics_",yesterday,".xlsm")

#pretty dates
today_display <- format(today, "%A, %b. %d, %Y")

download.file("https://www.pavoterservices.pa.gov/2024%20General%20Daily%20Mail%20Ballot%20Report.xlsx", filename) #download file

ballots <- read_excel(filename, sheet = "Total", trim_ws = TRUE, skip = 1, col_names = TRUE) #read file

```

```{r, echo=FALSE, warning=FALSE}

ballots_clean <- ballots %>% 
  filter(`CountyName` != 'TOTAL') %>%  #filter out only TOTAL row
  replace(is.na(.), "0") %>%  #replace everything that's "NULL" with a "0" 
  mutate(`Total Applications Approved`= as.numeric(`Total Applications Approved`),
         `Total Ballots Returned` = as.numeric(`Total Ballots Returned`), 
         `Democratic Applications Approved` = as.numeric(`Dem Applications Approved`), 
         `Republican Applications Approved` = as.numeric(`Rep Applications Approved`),
        `Other Applications Approved` = as.numeric(`Oth Applications Approved`),
        `Democratic Ballots Returned` = as.numeric(`Dem Ballots Returned`), 
        `Republican Ballots Returned` = as.numeric(`Rep Ballots Returned`), 
        `Other Ballots Returned` = as.numeric(`Oth Ballots Returned`)) %>% #convert them all to numbers
  rename(`Ballot Applications Approved`=`Total Applications Approved`,
         `Ballots Returned`=`Total Ballots Returned`,
         County = `CountyName`) %>% 
  mutate(`Percent Returned` = round(((`Ballots Returned`/`Ballot Applications Approved`)*100), digits=2)) #percent that have been returned

ballots_clean <- ballots_clean %>% select(`County`, `Ballot Applications Approved`, `Ballots Returned`, `Democratic Applications Approved`,`Republican Applications Approved`,`Other Applications Approved`,`Democratic Ballots Returned`,`Republican Ballots Returned`,`Other Ballots Returned`,`Percent Returned`)  # drop duplicate columns

write.csv(ballots_clean, "ballots_clean.csv", row.names = FALSE)

total_ballots_issued = sum(ballots_clean$`Ballot Applications Approved`)
total_ballots_returned = sum(ballots_clean$`Ballots Returned`)
total_pct_returned = round(((total_ballots_returned/total_ballots_issued)*100), digits=1)

if (total_ballots_issued > 1000000) {
  total_ballots_issued_display = paste(sep="", as.character(round((total_ballots_issued/1000000), digits=1)), " M")
} else {
  total_ballots_issued_display = paste(sep="", as.character(round((total_ballots_issued/1000), digits=1)), " K")
}


if (total_ballots_returned > 1000000) {
  total_ballots_returned_display = paste(sep="", as.character(round((total_ballots_returned/1000000), digits=1)), " M")
} else {
  total_ballots_returned_display = paste(sep="", as.character(round((total_ballots_returned/1000), digits=1)), " K")
}


```

<h1>Pennsylvania Mail Ballot Tracker</h1>
<h4>Last updated `r today_display` </h4>
<br>
<div class="cardContainer">
<div class="card" style="background-color:#00318b;">
<p>Ballots issued<br>
<span class="bignumber">`r total_ballots_issued_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>Ballots returned<br>
<span class="bignumber">`r total_ballots_returned_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>Percent returned<br>
<span class="bignumber">`r total_pct_returned`%<br></span><p>
</div>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

<iframe title="Share of mail ballots returned" aria-label="Map" id="datawrapper-chart-YxRx6" src="https://datawrapper.dwcdn.net/YxRx6/3/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="478" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(a){if(void 0!==a.data["datawrapper-height"]){var e=document.querySelectorAll("iframe");for(var t in a.data["datawrapper-height"])for(var r=0;r<e.length;r++)if(e[r].contentWindow===a.source){var i=a.data["datawrapper-height"][t]+"px";e[r].style.height=i}}}))}();
</script>

<br>
<br>
<br>
<br>


<iframe title="PA mail ballots by registered party" aria-label="Table" id="datawrapper-chart-l4o8R" src="https://datawrapper.dwcdn.net/l4o8R/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="461" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(a){if(void 0!==a.data["datawrapper-height"]){var e=document.querySelectorAll("iframe");for(var t in a.data["datawrapper-height"])for(var r=0;r<e.length;r++)if(e[r].contentWindow===a.source){var i=a.data["datawrapper-height"][t]+"px";e[r].style.height=i}}}))}();
</script>

<br>
<br>
<br>
<br>
<br>
<br>
<div class="footer">Built and designed by Maia Rosenfeld. Source: <a href="https://www.pa.gov/en/agencies/vote/elections/elections-data.html" target="_blank">Pennsylvania Department of State</a>. Data usually lags by about a day.</div>


```{r, echo=FALSE, warning=FALSE}

#trying to get date from header

#x <- XLConnect::loadWorkbook("vbm-statistics.xlsm")
#df <- XLConnect::readWorksheet(x, sheet=1)

#sheet1 <- x$worksheets[[1]]

#headers <- sheet1$headerFooter

```
